\documentclass[twoside,11pt]{article}

\usepackage{jmlr2e}
\usepackage{amsmath}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{enumitem}
\usepackage{lastpage}

\jmlrheading{0}{2026}{1-\pageref{LastPage}}{2/26}{}{PAPER-B-0001}{Jonathan Herrera-Vasquez}
\ShortHeadings{LIME vs SHAP: A Comprehensive Empirical Comparison}{Herrera-Vasquez}
\firstpageno{1}

\begin{document}

\title{LIME vs SHAP: A Comprehensive Empirical Comparison}

\author{
\name Jonathan Herrera-Vasquez \email jonnabio@gmail.com \\
\addr UNADE
}

\editor{To be assigned}

\maketitle

\begin{abstract}

LIME and SHAP are widely used explainability methods, yet guidance on
when to use each remains limited and often anecdotal. We present the
first comprehensive empirical comparison evaluating both methods across
six complementary metrics, using a rigorous benchmarking framework. Our
study on tabular data reveals strong, statistically robust directional
separation: SHAP dominates on stability and faithfulness-oriented
quality metrics, while LIME yields substantially sparser explanations
and lower latency in most matched contexts. The cost gap is large but
heterogeneous across model families, with a heavy-tailed SHAP runtime
distribution (matched-cell medians: 65.6 ms for SHAP versus 3.9 ms for
LIME; means: 18,104.5 ms versus 191.1 ms). Based on these findings, we
provide deployment-oriented guidance: SHAP for high-assurance auditing
and LIME for strict latency-oriented interfaces.
\end{abstract}

\begin{keywords}
explainable ai, lime, shap, interpretability, robustness benchmarking
\end{keywords}

\section{Introduction}

The ``LIME vs SHAP'' dilemma is a recurring theme in the deployment of
Explainable AI (XAI). Practitioners must often choose between the local
surrogate approach of LIME \citep{ribeiro2016why} and the game-theoretic
approach of SHAP \citep{lundberg2017unified}. While theoretical distinctions
are well-documented---LIME optimizes for local fidelity via sparse
linear models, while SHAP optimizes for Shapley value
properties---evaluation remains methodologically fragmented in practice.
Recent synthesis work reports that XAI still lacks a gold-standard
evaluation protocol and that evidence is often scattered across methods,
domains, and evaluation criteria \citep{longo2024manifesto}. The same
analysis also highlights robustness sensitivity in attribution methods
such as LIME and SHAP under perturbation settings, which makes
single-metric comparisons insufficient \citep{longo2024manifesto}.

This paper addresses the gap by providing a rigorous, multi-metric
comparison. We leverage the \textbf{XAI Evaluation Framework}
as an audited, configuration-driven benchmarking protocol to compare
these methods on tabular data across five key dimensions: Fidelity,
Stability, Sparsity, Computational Cost, and Faithfulness Gap.

Our contributions are:
\begin{enumerate}[leftmargin=*]
\item \textbf{Quantified Stability Gap}: We demonstrate a large stability
advantage for SHAP, quantifying the risk of LIME's nondeterminism.
\item \textbf{Cost-Fidelity Trade-off Analysis}: We map the Pareto frontier
of explanation quality versus latency.
\item \textbf{Semantic Evaluation}: We apply LLM-based judges to assess the
human interpretability of generated explanations.
\item \textbf{Practical Guidelines}: We synthesize these metrics into a
decision framework for practitioners.
\end{enumerate}

\section{Background}
This section summarizes the two explainability methods compared in our
benchmark and the properties most relevant to practitioners.

\subsection{LIME (Local Interpretable Model-Agnostic Explanations)}

LIME approximates a black-box model \(f\) locally around an instance
\(x\) by training an interpretable surrogate model \(g\) (e.g., linear
regression) on perturbed samples weighted by a kernel \(\pi_x\). It
solves:
\[
\xi(x) = \arg\min_{g \in G} \mathcal{L}(f, g, \pi_x) + \Omega(g).
\]
where \(\Omega(g)\) penalizes complexity (sparsity). LIME is favored for
its speed and explicit sparsity controls.

\subsection{SHAP (Shapley Additive Explanations)}

SHAP attributes prediction output to features based on their marginal
contribution across all possible coalitions, satisfying axioms of
efficiency, symmetry, and dummy properties. KernelSHAP is a
model-agnostic approximation of these values. SHAP is favored for its
theoretical grounding and consistency guarantees.

\section{Methodology}
This paper uses the same audited experimental stack developed for the
thesis framework and narrows it to a focused LIME-versus-SHAP
comparison. The goal is to produce claim-ready evidence on quality-cost
trade-offs under controlled, reproducible conditions.

\subsection{Research Questions and Hypotheses}
The methodology is aligned with the thesis objective of building a
model-agnostic, reproducible XAI evaluation protocol and validating it
with both quantitative and semantic criteria.

We address the following questions:
\begin{enumerate}[leftmargin=*]
\item \textbf{RQ1 (Robustness)}: Is SHAP more stable than LIME under
input perturbation?
\item \textbf{RQ2 (Parsimony)}: Is LIME sparser than SHAP?
\item \textbf{RQ3 (Efficiency)}: Is LIME computationally cheaper than
SHAP?
\item \textbf{RQ4 (Faithfulness)}: Does SHAP provide stronger
faithfulness/fidelity signals than LIME?
\end{enumerate}

Corresponding directional hypotheses are:
\begin{enumerate}[leftmargin=*]
\item \(H_1: S_{\mathrm{SHAP}} > S_{\mathrm{LIME}}\) (stability).
\item \(H_2: P_{\mathrm{LIME}} > P_{\mathrm{SHAP}}\) (sparsity).
\item \(H_3: C_{\mathrm{LIME}} < C_{\mathrm{SHAP}}\) (cost).
\item \(H_4: F_{\mathrm{SHAP}} > F_{\mathrm{LIME}}\) (fidelity/faithfulness).
\end{enumerate}

\subsection{Experimental Design and Analysis Units}
The robustness cohort is instantiated from the repository-tracked
EXP2-scaled matrix with factors:
\begin{itemize}[leftmargin=*]
\item model family \(g \in \{\mathrm{logreg}, \mathrm{rf}, \mathrm{xgb},
\mathrm{svm}, \mathrm{mlp}\}\),
\item explainer \(k \in \{\mathrm{lime}, \mathrm{shap}\}\),
\item seed \(s \in \{42, 123, 456, 789, 999\}\),
\item sampling intensity \(N \in \{50, 100, 200\}\) per TP/TN/FP/FN stratum.
\end{itemize}

This defines \(5 \times 2 \times 5 \times 3 = 150\) planned LIME/SHAP
runs. In the current artifact state, we retain 140 analyzable LIME/SHAP
runs (75 LIME, 65 SHAP) and 65 matched SHAP-LIME cells on identical
\((g,s,N)\) coordinates for paired inference.

To avoid pseudoreplication, we use a hierarchical analysis structure:
\begin{enumerate}[leftmargin=*]
\item \textbf{Instance level}: per-case metric values.
\item \textbf{Run level}: mean metric per configuration.
\item \textbf{Matched-cell level}: paired SHAP-LIME run means on shared
\((g,s,N)\) cells (primary inferential unit).
\end{enumerate}

\subsection{Data, Preprocessing, and Model Controls}
All experiments use the UCI Adult data set \citep{kohavi1996scaling} with a
stratified split and deterministic seeds. The preprocessing pipeline
follows train-only fitting with:
\begin{itemize}[leftmargin=*]
\item numeric scaling (`StandardScaler`),
\item categorical one-hot encoding (`OneHotEncoder` with
\texttt{handle\_unknown='ignore'}).
\end{itemize}

Pre-trained black-box models are loaded from fixed artifacts and
reused across explainers to isolate explainer effects:
random forest \citep{breiman2001random} and gradient-boosted trees
(\texttt{xgb}) \citep{chen2016xgboost}, alongside additional tabular
families present in the robustness cohort.

Evaluation instances are sampled by error quadrants (TP, TN, FP, FN) via
stratified sampling. Nominal run size is \(4N\), with realized counts
bounded by quadrant availability and runtime failures (observed range:
27 to 800 instances; median 400).

\subsection{Explainer Protocol}
Both explainers consume the same transformed feature space and target the
positive class probability:
\begin{itemize}[leftmargin=*]
\item \textbf{SHAP}: TreeExplainer for tree models, KernelExplainer for
non-tree models, with background subsampling (\(n=50\)).
\item \textbf{LIME}: tabular local surrogate with
\texttt{num\_features=10}, effective \texttt{num\_samples=1000}, and
\texttt{kernel\_width=3.0} in the EXP2 configuration family.
\end{itemize}

\subsection{Metric Operationalization}
Primary metrics are computed per instance and aggregated per run:
\begin{enumerate}[leftmargin=*]
\item \textbf{Stability} \(S\): mean pairwise cosine similarity of
attribution vectors under Gaussian perturbations (\(T=15\),
\(\sigma=0.1\)):
\[
S = \frac{2}{T(T-1)}\sum_{a<b}\cos\!\left(e^{(a)},e^{(b)}\right).
\]
\item \textbf{Sparsity} \(P\): active-feature ratio above threshold
\(\tau=10^{-4}\) (lower active ratio indicates sparser explanations).
\item \textbf{Fidelity} \(F\): faithfulness correlation between
attribution magnitude and prediction drop under feature masking.
\item \textbf{Faithfulness Gap} \(\Delta_k\): absolute prediction change
after masking top-\(k\) features (\(k=5\)):
\[
\Delta_k=\left|f(x)-f(x_{\mathrm{mask\text{-}top\text{-}k})}\right|.
\]
\item \textbf{Cost} \(C\): per-instance wall-clock explanation time (ms).
\end{enumerate}

\subsection{Inferential Protocol}
The statistical plan follows established comparative-ML practice
\citep{demsar2006statistical} and thesis ADR decisions:
\begin{enumerate}[leftmargin=*]
\item primary test: two-sided paired Wilcoxon signed-rank on matched
SHAP-LIME cells for each metric \citep{wilcoxon1945individual},
\item effect size: paired \(d_z\) for magnitude reporting
\citep{lakens2013calculating},
\item multiplicity control: Holm-Bonferroni over the five primary
metrics \citep{holm1979simple}.
\end{enumerate}

All inferential claims are restricted to artifact-qualified runs:
parseable, non-empty results with explicit matched pairing.

\subsection{Secondary Semantic Evaluation}
Consistent with thesis objectives, the framework includes an LLM-based
semantic layer (clarity/intuitiveness-style scoring). In this paper,
semantic evaluation is treated as secondary evidence and does not replace
technical-metric inference. Its role is to contextualize technical
trade-offs, not to act as the primary statistical endpoint.

\subsection{Reproducibility Contract}
The protocol is fully configuration-driven (YAML + seeded execution),
with per-run `results.json` artifacts, explicit exclusion of malformed or
empty runs, and no synthetic imputation. This provides a deterministic
audit trail from configuration to manuscript-level claim.

\section{Results}
Results are reported from the artifact-qualified EXP2 robustness cohort.
Of 250 run files, 233 are analyzable; for this paper's paired
LIME-versus-SHAP inference we retain 140 analyzable runs (75 LIME, 65
SHAP) and 65 matched cells with identical \((\mathrm{model}, \mathrm{seed},
N)\). Matched coverage spans all five model families
(\(\mathrm{xgb}=15\), \(\mathrm{logreg}=15\), \(\mathrm{rf}=15\),
\(\mathrm{mlp}=11\), \(\mathrm{svm}=9\)) and all sampling intensities
\(N\in\{50,100,200\}\) with near-balanced counts.

\subsection{Primary Paired Inference}
Table~\ref{tab:paired_main} summarizes paired run-level outcomes on the
65 matched cells. All five primary endpoints remain significant under
Holm-adjusted testing.

\begin{table}[t]
\centering
\caption{Matched SHAP-versus-LIME paired results (\(n=65\) cells). For all metrics,
\(\Delta=\) SHAP - LIME. Sparsity is measured as active-feature ratio, so larger values indicate denser explanations.}
\label{tab:paired_main}
\begin{tabular}{lrrrrr}
\toprule
Metric & LIME Mean & SHAP Mean & \(\Delta\) & \(p\) (Wilcoxon) & \(d_z\) \\
\midrule
Stability & 0.014 & 0.742 & +0.728 & \(2.4\times10^{-12}\) & 3.07 \\
Sparsity (active ratio) & 0.085 & 0.247 & +0.162 & \(2.4\times10^{-12}\) & 0.94 \\
Fidelity & 0.558 & 0.812 & +0.254 & \(2.4\times10^{-12}\) & 3.97 \\
Faithfulness Gap & 0.336 & 0.429 & +0.093 & \(2.4\times10^{-12}\) & 0.94 \\
Cost (ms) & 191.1 & 18104.5 & +17913.4 & \(3.4\times10^{-9}\) & 0.36 \\
\bottomrule
\end{tabular}
\end{table}

The stability effect is the largest practical separation in the study.
Median paired stability difference is \(+0.864\) in favor of SHAP, and
the direction is consistent in all 65 matched cells. This indicates that
LIME's local surrogate sampling introduces substantial run-to-run
volatility under perturbation, whereas SHAP is comparatively invariant.

For sparsity, the sign is also consistent in all 65 matched cells, but
interpretation requires care: the metric here is active-feature ratio.
SHAP's larger value means denser explanations; thus LIME is systematically
sparser and therefore easier to compress for user-facing narratives.

Fidelity and faithfulness gap both favor SHAP in all matched cells. The
fidelity margin (\(+0.254\) mean paired difference) and very large
paired effect size (\(d_z=3.97\)) indicate not only statistical
separation but practically meaningful quality gains under this protocol.

\subsection{Runtime Distribution and Heterogeneity}
The cost signal is significant but strongly heterogeneous. SHAP is slower
in 49 of 65 matched cells, with a median cost ratio of approximately
\(9.9\times\) (SHAP/LIME). However, the distribution is heavy-tailed:
for SHAP, the 90th and 95th percentile latencies are approximately
89,261 ms and 95,781 ms, respectively, compared with 798 ms and
817 ms for LIME. This tail behavior is concentrated in kernel-SHAP
contexts (notably SVM), which inflates mean cost beyond median behavior.

Heterogeneity analysis by model family shows that quality ordering
(stability, sparsity-direction, fidelity, faithfulness gap) is robust,
while cost ranking is model dependent. In particular, all 15 XGBoost
matched cells show SHAP faster than LIME, whereas RF, SVM, MLP, and
LogReg are predominantly or uniformly SHAP-slower. Therefore, ``LIME is
faster'' is valid as a majority statement in this cohort, but it is not
universally true across model classes.

\subsection{Synthesis of Trade-offs}
The empirical frontier is clear: SHAP delivers higher reliability and
faithfulness at substantially higher expected runtime, while LIME delivers
lower latency and stronger compression at the cost of robustness. The
results therefore support objective-driven selection rather than global
method supremacy. For assurance-critical analysis, SHAP's quality margin
is large and consistent. For interaction-critical systems, LIME's runtime
profile remains operationally attractive.

\section{Practical Recommendations}

Based on these empirical findings, we propose the \textbf{Hybrid
Deployment Pattern}:

\begin{enumerate}[leftmargin=*]
\item \textbf{Tier 1: Real-Time / End-User (Use LIME)}
\begin{itemize}[leftmargin=*]
\item Use when explaining predictions to end-users in a live application
(e.g., ``Why was my loan denied?'' in a mobile app).
\item Requirement: \(< 200\) ms latency.
\item Acceptance criterion: explanations are ``generally correct'' sketches.
Reduced fidelity is an acceptable trade-off for usability.
\end{itemize}
\item \textbf{Tier 2: Auditing / Data Science (Use SHAP)}
\begin{itemize}[leftmargin=*]
\item Use when analyzing model bias, debugging model logic, or responding
to regulatory inquiries.
\item Requirement: exactness and stability.
\item Acceptance criterion: latency of seconds or minutes is acceptable.
Evaluating a batch of data can be done offline.
\end{itemize}
\end{enumerate}

\section{Limitations and Future Work}

\begin{itemize}
\item \textbf{Data Set Scope}: This study focuses on tabular data (Adult).
Image and text domains may exhibit different trade-offs.
\item \textbf{Configuration Space}: We compared standard configurations
(KernelSHAP versus standard LIME). Hyperparameter tuning (e.g., LIME
kernel width) can alter results.
\item \textbf{Adversarial Robustness}: We did not subject the methods to
active adversarial attacks, only random perturbations.
\end{itemize}

\section{Conclusion}

The choice between LIME and SHAP is not a matter of ``better,'' but of
``fit for purpose.'' SHAP is the scientifically rigorous choice for
understanding \emph{what the model actually does}, while LIME is the
pragmatic choice for \emph{explaining it quickly}. Our paired analysis
across 65 matched robustness cells shows consistent SHAP advantages in
stability and faithfulness-oriented quality, alongside strong LIME
advantages in sparsity and latency for most (but not all) model
contexts. This evidence supports architecture-level deployment patterns
that separate user-facing low-latency explanation paths from
high-assurance auditing workflows.

\acks{No external funding supported this work. The author declares no
competing interests.}

\clearpage
\phantomsection
\pdfbookmark[1]{References}{references}
\begin{thebibliography}{99}

\bibitem[Breiman(2001)]{breiman2001random}
Breiman, L. (2001).
Random forests.
\emph{Machine Learning}, 45(1):5--32.

\bibitem[Chen and Guestrin(2016)]{chen2016xgboost}
Chen, T. and Guestrin, C. (2016).
XGBoost: A scalable tree boosting system.
In \emph{Proceedings of the 22nd ACM SIGKDD International Conference on
Knowledge Discovery and Data Mining}, 785--794.

\bibitem[Dems\v{s}ar(2006)]{demsar2006statistical}
Dems\v{s}ar, J. (2006).
Statistical comparisons of classifiers over multiple data sets.
\emph{Journal of Machine Learning Research}, 7:1--30.

\bibitem[Holm(1979)]{holm1979simple}
Holm, S. (1979).
A simple sequentially rejective multiple test procedure.
\emph{Scandinavian Journal of Statistics}, 6(2):65--70.

\bibitem[Kohavi(1996)]{kohavi1996scaling}
Kohavi, R. (1996).
Scaling up the accuracy of Naive-Bayes classifiers: A decision-tree
hybrid.
In \emph{Proceedings of the Second International Conference on Knowledge
Discovery and Data Mining}, 202--207.

\bibitem[Lakens(2013)]{lakens2013calculating}
Lakens, D. (2013).
Calculating and reporting effect sizes to facilitate cumulative science:
A practical primer for \(t\)-tests and ANOVAs.
\emph{Frontiers in Psychology}, 4:863.

\bibitem[Longo et~al.(2024)Longo, Brcic, Cabitza, Choi, Confalonieri, Del
Ser, Guidotti, Hayashi, Herrera, Holzinger, Jiang, Khosravi, Lecue,
Malgieri, P\'{a}ez, Samek, Schneider, Speith, and Stumpf]{longo2024manifesto}
Longo, L., Brcic, M., Cabitza, F., Choi, J., Confalonieri, R., Del Ser, J.,
Guidotti, R., Hayashi, Y., Herrera, F., Holzinger, A., Jiang, R.,
Khosravi, H., Lecue, F., Malgieri, G., P\'{a}ez, A., Samek, W.,
Schneider, J., Speith, T., and Stumpf, S. (2024).
Explainable artificial intelligence (XAI) 2.0: A manifesto of open
challenges and interdisciplinary research directions.
\emph{Information Fusion}, 106:102301.
\url{https://doi.org/10.1016/j.inffus.2024.102301}.

\bibitem[Lundberg and Lee(2017)Lundberg and Lee]{lundberg2017unified}
Lundberg, S. M. and Lee, S.-I. (2017).
A unified approach to interpreting model predictions.
In \emph{Advances in Neural Information Processing Systems}, 4765--4774.

\bibitem[Ribeiro et~al.(2016)Ribeiro, Singh, and Guestrin]{ribeiro2016why}
Ribeiro, M. T., Singh, S., and Guestrin, C. (2016).
Why should I trust you?: Explaining the predictions of any classifier.
In \emph{Proceedings of the 22nd ACM SIGKDD International Conference on
Knowledge Discovery and Data Mining}, 1135--1144.

\bibitem[Wilcoxon(1945)]{wilcoxon1945individual}
Wilcoxon, F. (1945).
Individual comparisons by ranking methods.
\emph{Biometrics Bulletin}, 1(6):80--83.

\end{thebibliography}

\end{document}
