\documentclass[twoside,11pt]{article}

\usepackage{jmlr2e}
\usepackage{amsmath}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{enumitem}
\usepackage{lastpage}

\jmlrheading{0}{2026}{1-\pageref{LastPage}}{Submitted 2/2026}{Under Review}{PAPER-A-0001}{Jonathan Herrera-Vasquez}
\ShortHeadings{XAI Benchmarking Framework and Operation Protocol}{XAI Benchmarking Framework}
\firstpageno{1}

\begin{document}

\title{A Framework for Rigorous Evaluation of Model-Agnostic Explainability Methods:\\
Multi-Metric Statistical Benchmarking, Operational Protocol, and Reproducibility}

\author{
\name Jonathan Herrera-Vasquez \email jonnabio@gmail.com \\
\addr UNADE
}

\editor{TBD}

\maketitle

\begin{abstract}
Evaluating explainability methods requires more than a single faithfulness proxy. We present a modular benchmarking framework centered on quantitative XAI quality metrics: fidelity, stability, sparsity, computational cost, and faithfulness gap, plus an explicit method for operating the framework end-to-end. On the UCI Adult benchmark, we use a staged protocol (EXP1 calibration/reproducibility and EXP2 comparative/robustness benchmarking); the robustness cohort currently contains 250 of 300 planned configurations (83.3\% coverage). Across complete model-size blocks ($5$ models, $N \in \{50,100,200\}$), Friedman tests indicate significant method differences for fidelity ($\chi^2=42.12$, $p=3.78\times10^{-9}$), stability ($\chi^2=43.88$, $p=1.60\times10^{-9}$), sparsity ($\chi^2=35.64$, $p=8.92\times10^{-8}$), faithfulness gap ($\chi^2=45.00$, $p=9.25\times10^{-10}$), and runtime ($\chi^2=27.72$, $p=4.16\times10^{-6}$). SHAP leads on fidelity/stability, DiCE leads on sparsity, and LIME is generally fastest and most practical outside SVM-KernelSHAP bottlenecks. We release the framework, operation protocol, and artifacts with explicit data-quality caveats for reproducible benchmark use. LLM-based semantic scoring is deferred to a dedicated follow-up study.
\end{abstract}

\begin{keywords}
Explainable AI, benchmark methodology, reproducibility, statistical evaluation, model-agnostic explanations
\end{keywords}

\section{Introduction}
Evaluation of post-hoc explainability methods remains methodologically fragmented.
Recent review evidence shows that XAI studies use heterogeneous metrics and
often report inconsistent evaluation objectives across domains, which limits
comparability and cumulative progress \citep{kadir2023metrics,pawlicki2024multiple}.
In practice, fidelity-centric reporting is still common, even when stability,
complexity, and computational burden materially affect whether explanations are
usable in decision pipelines.

Open-source tooling has improved reproducibility at the implementation layer,
including interfaces for large sets of explainers and metrics
\citep{li2022interpretdl,hedstrom2023quantus}. However, tool-level coverage
does not by itself establish inferential comparability: benchmark claims still
require explicit controls for split discipline, artifact qualification,
aggregation units, and multiplicity-aware statistical testing.

Parallel work in graph explainability benchmarking reinforces this point.
GraphXAI, ShapeGGen, and B-XAIC show that benchmark value depends on explicit
dataset assumptions, ground-truth design choices, and transparent reporting of
evaluation constraints \citep{agarwal2023gnn_eval,proszewska2025bxaic}. At the
same time, recent findings on attribution consensus in Rashomon sets indicate
that explanation variability can persist even under similar predictive
performance \citep{laberge2023rashomon}. Faithfulness-oriented studies also
show that metric behavior is sensitive to perturbation design and protocol
choices, requiring robustness-aware interpretation \citep{zheng2025ffidelity}.

This paper focuses on model-agnostic, tabular post-hoc explanation benchmarking.
It does not target representation-level interpretability (e.g., neuron-level
analysis in NLP models), which has distinct constructs and evaluation
methodologies \citep{durrani2023salient}.

Given these gaps, we operationalize a benchmark-first framework that combines
multi-metric scoring with an auditable operation protocol:
\begin{enumerate}[leftmargin=*]
    \item A generic trainer/explainer architecture for model-agnostic evaluation.
    \item A prescriptive operation method that defines how the framework is run, audited, and reported.
    \item Multi-metric quantitative scoring with explicit trade-off analysis.
    \item Reproducibility and statistical rigor suitable for benchmark publication.
\end{enumerate}

\section{Related Work and Framework Design}

\subsection{Literature-Grounded Rationale}
The literature motivating this design supports five requirements:
\begin{enumerate}[leftmargin=*]
    \item Multi-metric evaluation is necessary.
    \item Faithfulness is multi-faceted.
    \item Computational tractability must be reported explicitly.
    \item Benchmark transparency and artifact accounting are first-order requirements.
    \item LLM-judge scoring should be treated as a separate calibrated track.
\end{enumerate}

\subsection{Architecture}
The framework decouples model training, explanation generation, and evaluation through a registry/factory design. Current model coverage includes \texttt{logreg}, \texttt{rf}, \texttt{xgb}, \texttt{svm}, and \texttt{mlp}, with \texttt{lime}, \texttt{shap}, \texttt{anchors}, and \texttt{dice} explainers.

\subsection{Metric Suite (Primary)}
Primary quantitative metrics are:
\begin{itemize}[leftmargin=*]
    \item Fidelity
    \item Stability
    \item Sparsity
    \item Cost
    \item Faithfulness gap
\end{itemize}

\subsection{Paper A Operational Outcome: Framework Operation Method (FOM-7)}
Paper A contributes not only benchmark findings, but also a reusable operating method for end-to-end benchmark execution, auditing, and claim formation.

\section{Methodology}

\subsection{Study Scope and Evidence Cohorts}
\label{sec:methods-scope}
To separate development from confirmation, Paper A uses staged evidence cohorts
with explicit claim roles.
\begin{table}[t]
\centering
\caption{Evidence cohorts and their role in Paper A claims.}
\begin{tabularx}{\linewidth}{l X X}
\toprule
\textbf{Cohort} & \textbf{Primary objective} & \textbf{Role in claims} \\
\midrule
EXP1 pilot/calibration & Implementation sanity checks before scaled runs. & Not used for confirmatory significance claims. \\
EXP1 reproducibility & Repeated-seed variance profiling on core settings. & Supports reproducibility characterization (dispersion/CV). \\
EXP2 comparative & Broad method-model coverage under fixed seeds. & Descriptive comparative context. \\
EXP2 robustness & Multi-seed, multi-sample-size benchmark runs. & Primary inferential cohort for global and paired tests. \\
\bottomrule
\end{tabularx}
\label{tab:cohorts}
\end{table}

\subsection{Research Questions, Factors, and Analysis Units}
\label{sec:methods-design}
The confirmatory questions are:
\begin{enumerate}[leftmargin=*]
    \item \textbf{RQ1 (global separation):} do explainer methods differ on the primary metrics under matched contexts?
    \item \textbf{RQ2 (targeted pairwise contrast):} on matched cells, does SHAP differ from LIME on quality-cost trade-offs?
    \item \textbf{RQ3 (reproducibility profile):} how much metric dispersion appears under repeated runs?
\end{enumerate}

EXP2 is a crossed design with model family $g$, explainer method $k$, seed $s$,
and per-quadrant sampling intensity $n$:
\[
g\in\{\text{logreg},\text{rf},\text{xgb},\text{svm},\text{mlp}\},\quad
k\in\{\text{shap},\text{lime},\text{anchors},\text{dice}\},
\]
\[
s\in\{42,123,456,789,999\},\quad
n\in\{50,100,200\}.
\]
Planned EXP2 robustness size is $5\times4\times5\times3=300$ runs. In the
current snapshot, 250 artifacts are present and 233 are analyzable (16 empty,
1 malformed).

The inferential unit is not the instance row. For metric $m$, the run-level
estimand is $y_{g,k,s,n,m}$ and the matched block is $b=(g,n)$. Tests are
performed on run/block aggregates to avoid pseudo-replication.

\subsection{Data Protocol and Leakage Controls}
\label{sec:methods-data}
The benchmark uses Adult Income loaded through
\path{src/data_loading/adult.py} with OpenML as primary source and UCI fallback.
Cleaning applies whitespace normalization, missing-value canonicalization,
target normalization, duplicate removal, and deterministic feature ordering.

Leakage controls are fixed across all explainers:
\begin{enumerate}[leftmargin=*]
    \item stratified split before any preprocessing fit,
    \item train-only preprocessor fitting,
    \item shared persisted preprocessor across methods in each run,
    \item explanation on the same transformed feature space used by the model.
\end{enumerate}

Data split is
\[
\texttt{train\_test\_split(test\_size=0.2,\ stratify=y,\ random\_state=seed)}.
\]
For seed 42, the split is 39,032 train and 9,758 test rows. Each run samples
TP/TN/FP/FN strata with target $N$ per stratum (nominal run size $4N$, subject
to stratum capacity).

\subsection{Model and Explanation Protocol}
\label{sec:methods-protocol}
EXP2 evaluates frozen EXP1 model artifacts; models are not retrained during
EXP2. Model families are \texttt{logreg}, \texttt{rf}, \texttt{xgb},
\texttt{svm}, and \texttt{mlp}, with fixed repository-tracked settings.

All methods are executed through
\texttt{ExplainerWrapper.explain\_instance} and timed per instance.
Effective explainer conditions are:
\begin{enumerate}[leftmargin=*]
    \item SHAP: tree mode for RF and XGB; kernel mode for LogReg, SVM, and MLP;
    background sample count set to 50.
    \item LIME: \texttt{num\_samples=1000}, \texttt{num\_features=10},
    \texttt{kernel\_width=3.0}.
    \item Anchors: \texttt{AnchorTabular} on transformed training data with
    effective threshold fixed at 0.95.
    \item DiCE: random backend, opposite-class counterfactual target, importance
    from absolute original--counterfactual deltas.
\end{enumerate}

Two implementation-bound caveats are explicit in claim interpretation:
Anchors YAML threshold is not consumed at runtime (fixed 0.95), and DiCE YAML
counterfactual count is not consumed (effective \texttt{total\_CFs=1}).

\subsection{Metrics and Statistical Inference}
\label{sec:methods-inference}
Primary metrics are fidelity, stability, sparsity, faithfulness gap, and cost
(milliseconds). Aggregation is instance $\rightarrow$ run $\rightarrow$ block.
EEU is not used for confirmatory inference in this draft.

Faithfulness and stability are computed as:
\begin{align}
\Delta_k &= \left|f(x)-f\left(x_{\text{mask-top-}k}\right)\right|,\quad
\rho = \mathrm{corr}\left(|w_i|,\left|f(x)-f(x_{-i})\right|\right),\\
x^{(t)} &= x + \epsilon^{(t)},\ \epsilon^{(t)}\sim\mathcal{N}(0,\sigma^2),\quad
S = \frac{2}{T(T-1)}\sum_{a<b}\cos\left(e^{(a)},e^{(b)}\right).
\end{align}

Inferential workflow:
\begin{enumerate}[leftmargin=*]
    \item Friedman omnibus test per metric on complete $(g,n)$ blocks.
    \item Nemenyi post-hoc localization when Friedman is significant.
    \item Two-sided paired Wilcoxon SHAP--LIME tests on matched
    $(g,s,n)$ cells (45-cell primary set; 65-cell sensitivity set).
\end{enumerate}

Multiplicity handling uses Nemenyi within each metric and Holm-Bonferroni
across the five primary metrics for each inferential family (Friedman,
Wilcoxon-45, Wilcoxon-65). Reported uncertainty/effect summaries include
mean, standard deviation, CV, confidence intervals (where generated),
Kendall's $W$, Cohen's $d_z$, and median paired differences.

Artifact qualification is deterministic: malformed or empty runs are excluded,
no synthetic reconstruction is used, and omnibus tests are restricted to
block-complete contexts.

\subsection{Reproducibility, FOM-7, and Validity Boundaries}
\label{sec:methods-repro}
Paper A operationalizes the benchmark using FOM-7, a stage-gated protocol:
\begin{enumerate}[leftmargin=*]
    \item protocol freeze,
    \item controlled batch execution,
    \item artifact integrity audit,
    \item harmonization to analysis-ready tables,
    \item inferential export,
    \item reproducibility profiling,
    \item claim-ready reporting.
\end{enumerate}
Progression requires gate satisfaction at each stage; failed gates downgrade
affected outputs to descriptive-only status.

The reproducibility package includes code, experiment configurations, model and
preprocessor artifacts, per-run outputs, analysis exports, and figure scripts.
For concision, command-level rerun instructions, provenance hashes, and
environment anchors are maintained in repository manifests and scripts rather
than duplicated in the main text.

Validity controls and residual risks are summarized as follows:
\begin{enumerate}[leftmargin=*]
    \item \textbf{Internal validity:} explicit missingness audit and
    block-complete inference; residual risk from non-random coverage gaps.
    \item \textbf{Construct validity:} metric definitions tied to code-level
    implementations; residual risk from omitted secondary metrics in claims.
    \item \textbf{Statistical conclusion validity:} matched-design non-parametric
    tests with multiplicity control; residual risk from absent equivalence bands.
    \item \textbf{External validity:} diversity across model families and seeds,
    but single tabular dataset scope.
    \item \textbf{Implementation validity:} effective runtime parameter bindings
    are disclosed; harmonization changes may shift absolute levels.
\end{enumerate}

Scope of claims: conclusions in Section 4 are comparative and bounded to this
benchmark setting (Adult tabular task, declared models/explainers, and
artifact-qualified EXP2 subsets).

\section{Results}

\subsection{Overview of Evaluation Setup}
This section reports results from two evidence streams: EXP2 robustness (global
and paired method comparisons) and EXP1 reproducibility (variance profiling).
Compared explainers are SHAP, LIME, Anchors, and DiCE. Reported metrics are
Fidelity, Stability, Sparsity, Faithfulness Gap, and Cost (ms).
Metric definitions and the inferential workflow are specified in
Section~\ref{sec:methods-inference}.
``15 complete model-size blocks'' denotes the block-level EXP2 units used for
global inference, and ``45 matched configs'' denotes the SHAP--LIME subset with
shared $(\text{model},\text{seed},N)$ cells on \texttt{logreg/rf/xgb}.

\subsection{Global Trade-offs Across Methods (Quality vs Cost)}
\begin{table}[h]
\centering
\caption{Global method means (EXP2 robustness: 15 complete model-size blocks).}
\begin{tabular}{lrrrrr}
\toprule
Method & Fidelity & Stability & Sparsity & Faithfulness Gap & Time (ms) \\
\midrule
SHAP    & 0.8176 & 0.7377 & 0.2264 & 0.4474 & 685220.23 \\
LIME    & 0.5602 & 0.0144 & 0.0846 & 0.3342 & 3660.68 \\
Anchors & 0.3853 & 0.0006 & 0.0928 & 0.2382 & 25326.92 \\
DiCE    & 0.1716 & 0.3602 & 0.0164 & 0.0988 & 16306.50 \\
\bottomrule
\end{tabular}
\label{tab:global-tradeoffs}
\end{table}

SHAP has the highest Fidelity (0.8176), highest Stability (0.7377), and highest
Faithfulness Gap (0.4474). DiCE has the lowest Sparsity value (0.0164), while
LIME has the lowest Cost (3660.68 ms). Anchors shows very low Stability (0.0006)
with Cost (25326.92 ms) above LIME and DiCE.

The reported means show an explicit quality--cost frontier: SHAP occupies the
strongest quality region on Fidelity/Stability/Faithfulness Gap with the highest
runtime (685220.23 ms), whereas LIME occupies the lowest-cost region with lower
Fidelity and Stability. DiCE provides the lowest Sparsity value with lower
Fidelity and Faithfulness Gap.

\subsection{Global Statistical Evidence Across Methods}
\begin{table}[h]
\centering
\caption{Friedman tests across methods.}
\begin{tabular}{lrrl}
\toprule
Metric & Statistic & p-value & Significant \\
\midrule
Fidelity         & 42.12 & 3.78e-09 & Yes \\
Stability        & 43.88 & 1.60e-09 & Yes \\
Sparsity         & 35.64 & 8.92e-08 & Yes \\
Faithfulness Gap & 45.00 & 9.25e-10 & Yes \\
Cost             & 27.72 & 4.16e-06 & Yes \\
\bottomrule
\end{tabular}
\label{tab:friedman}
\end{table}

For every metric, the Friedman test rejects the null of equal performance across
methods. This establishes global between-method differences, but does not
identify pairwise direction without post-hoc localization.
Post-hoc localization uses Nemenyi pairwise comparisons within each metric,
with Holm-Bonferroni adjustment across the five primary metrics.

\subsection{Targeted Pairwise Comparison (SHAP vs LIME on Shared Families)}
\begin{table}[h]
\centering
\caption{Paired Wilcoxon comparison on 45 matched configurations (\texttt{logreg/rf/xgb}).}
\begin{tabular}{lrrr}
\toprule
Metric & SHAP Mean & LIME Mean & Wilcoxon p-value \\
\midrule
Fidelity         & 0.8112 & 0.5556 & 5.68e-14 \\
Stability        & 0.8013 & 0.0154 & 5.68e-14 \\
Sparsity         & 0.3156 & 0.0845 & 5.68e-14 \\
Faithfulness Gap & 0.3924 & 0.3408 & 5.68e-14 \\
Cost (ms)        & 1258.72 & 210.45 & 2.29e-06 \\
\bottomrule
\end{tabular}
\label{tab:shap-lime}
\end{table}

Within this matched subset, SHAP is higher on Fidelity, Stability, and
Faithfulness Gap, while LIME has lower Cost and lower Sparsity value. The
reported p-values indicate non-zero paired differences for all listed metrics.
Practically, the quality gains for SHAP are accompanied by a runtime penalty
relative to LIME under shared cells.

This comparison is restricted to \texttt{logreg/rf/xgb} matched configurations
and is not directly generalizable to non-matched families or to methods outside
the SHAP--LIME pair.

\subsection{Reproducibility and Variance Profile}
EXP1 repeated-run evidence indicates lower dispersion for quality-oriented
metrics and higher dispersion for runtime behavior. Quality metrics remain
relatively stable across repeated seeds, while Cost varies more strongly for
SHAP variants.
\begin{itemize}[leftmargin=*]
    \item Metric stability (quality): Fidelity CV $\leq 0.0634$, Stability CV $\leq 0.0826$, Sparsity CV $\leq 0.0441$, Faithfulness-gap CV $\leq 0.0358$.
    \item Runtime instability (cost): Cost CV reaches 0.225 for SHAP variants.
    \item Decision implication: quality rankings are more repeatable than runtime behavior under the reported repeated-run protocol.
\end{itemize}

\subsection{Synthesis and Implications for Practice}
\begin{itemize}[leftmargin=*]
    \item When the objective is maximizing Fidelity/Stability/Faithfulness Gap and compute budget is permissive, SHAP is the supported choice in this benchmark.
    \item When compute or latency is constrained, LIME is the supported low-cost option, with lower Fidelity and Stability than SHAP.
    \item When conciseness (Sparsity) is the primary objective, DiCE provides the lowest reported Sparsity value, with lower Fidelity and Faithfulness Gap.
    \item Methods are not statistically interchangeable: Friedman tests reject the null of equal performance across methods for all five primary metrics.
    \item On the shared SHAP--LIME subset, quality improvements for SHAP are paired with higher runtime cost.
\end{itemize}
Interpretation limits:
\begin{itemize}[leftmargin=*]
    \item Anchors stability values near zero should be interpreted with implementation caveats: effective runtime behavior uses fixed precision threshold 0.95 and binary rule-membership importance vectors, which can reduce perturbation-based similarity under this protocol.
    \item Method rankings are currently validated only on one tabular benchmark (Adult); external generalization requires replication on additional datasets and modalities before cross-domain claims are made.
\end{itemize}

\section{Conclusion}
The reported metrics indicate a multi-objective frontier rather than a single
best method: SHAP is strongest on the reported quality-oriented metrics, LIME is
strongest on runtime efficiency, and DiCE is strongest on the reported sparsity
criterion. Global tests reject the null of equal method performance across the
primary metrics, and paired SHAP--LIME analysis indicates a quality--cost
trade-off on shared model families.

Methodologically, Paper A contributes FOM-7 as an auditable benchmark operation
protocol that connects protocol freezing, controlled execution, integrity audit,
deterministic inference export, and claim-traceable reporting. This contribution
is operational and reproducibility-focused rather than algorithmic.

\begin{itemize}[leftmargin=*]
    \item \textbf{Takeaway:} method choice should be objective-driven (quality, sparsity, or runtime), not based on a single aggregated score.
    \item \textbf{Takeaway:} statistical evidence supports method differentiation at the omnibus level; pairwise direction must be interpreted within matched-design limits.
    \item \textbf{Limitation:} incomplete EXP2 coverage and malformed artifacts qualify confidence in full-grid frontier stability.
    \item \textbf{Next steps:} publish a missingness diagnostic by model/explainer/seed/sample-size and release a versioned reproducibility bundle with immutable DOI (code, configs, run outputs, inferential exports, and figure scripts).
    \item \textbf{Next steps:} integrate deferred semantic evaluation via a preregistered rubric, blinded human rating, inter-rater reliability reporting, and agreement analysis against the quantitative metrics.
\end{itemize}

\acks{Funding and competing-interest disclosures are to be finalized in the submission package. This draft was prepared from repository artifacts dated February 2026.}

\clearpage
\phantomsection
\pdfbookmark[1]{References}{references}
\begin{thebibliography}{10}

\bibitem[Agarwal et~al.(2023)Agarwal, Queen, Lakkaraju, and Zitnik]{agarwal2023gnn_eval}
Chirag Agarwal, Owen Queen, Himabindu Lakkaraju, and Marinka Zitnik.
2023.
Evaluating explainability for graph neural networks.
\emph{Scientific Data}, 10:144.

\bibitem[Durrani et~al.(2023)Durrani, Dalvi, and Sajjad]{durrani2023salient}
Nadir Durrani, Fahim Dalvi, and Hassan Sajjad.
2023.
Discovering Salient Neurons in deep NLP models.
\emph{Journal of Machine Learning Research}, 24:1--40.

\bibitem[Hedstr\"om et~al.(2023)Hedstr\"om, Weber, Bareeva, Krakowczyk, Motzkus, Samek, Lapuschkin, and H\"ohne]{hedstrom2023quantus}
Anna Hedstr\"om, Leander Weber, Dilyara Bareeva, Daniel Krakowczyk,
Franz Motzkus, Wojciech Samek, Sebastian Lapuschkin, and Marina M.-C. H\"ohne.
2023.
Quantus: An Explainable AI Toolkit for Responsible Evaluation of Neural Network Explanations and Beyond.
\emph{Journal of Machine Learning Research}, 24:1--11.

\bibitem[Kadir et~al.(2023)Kadir, Mosavi, and Sonntag]{kadir2023metrics}
Md Abdul Kadir, Amir Mosavi, and Daniel Sonntag.
2023.
Evaluation Metrics for XAI: A Review, Taxonomy, and Practical Applications.
In \emph{Proceedings of the 2023 IEEE 27th International Conference on
Intelligent Engineering Systems (INES)}, pages 111--124, Nairobi, Kenya.

\bibitem[Laberge et~al.(2023)Laberge, Pequignot, Mathieu, Khomh, and Marchand]{laberge2023rashomon}
Gabriel Laberge, Yann Pequignot, Alexandre Mathieu, Foutse Khomh, and Mario Marchand.
2023.
Partial Order in Chaos: Consensus on Feature Attributions in the Rashomon Set.
\emph{Journal of Machine Learning Research}, 24:1--50.

\bibitem[Li et~al.(2022)Li, Xiong, Li, Wu, Chen, and Dou]{li2022interpretdl}
Xuhong Li, Haoyi Xiong, Xingjian Li, Xuanyu Wu, Zeyu Chen, and Dejing Dou.
2022.
InterpretDL: Explaining Deep Models in PaddlePaddle.
\emph{Journal of Machine Learning Research}, 23:1--6.

\bibitem[Pawlicki et~al.(2024)Pawlicki, Pawlicka, Uccello, Szelest, D'Antonio, Kozik, and Chora\'s]{pawlicki2024multiple}
Marek Pawlicki, Aleksandra Pawlicka, Federica Uccello, Sebastian Szelest,
Salvatore D'Antonio, Rafa{\l} Kozik, and Micha{\l} Chora\'s.
2024.
Evaluating the necessity of the multiple metrics for assessing explainable AI: A critical examination.
\emph{Neurocomputing}, 602:128282.

\bibitem[Proszewska et~al.(2025)Proszewska, Danel, and Rymarczyk]{proszewska2025bxaic}
Magdalena Proszewska, Tomasz Danel, and Dawid Rymarczyk.
2025.
B-XAIC Dataset: Benchmarking Explainable AI for Graph Neural Networks Using Chemical Data.
\emph{arXiv preprint arXiv:2505.22252}.

\bibitem[Zheng et~al.(2025)Zheng, Shirani, Chen, Lin, Cheng, Guo, and Luo]{zheng2025ffidelity}
Xu Zheng, Farhad Shirani, Zhuomin Chen, Chaohao Lin, Wei Cheng, Wenbo Guo, and Dongsheng Luo.
2025.
F-FIDELITY: A Robust Framework for Faithfulness Evaluation of Explainable AI.
In \emph{International Conference on Learning Representations (ICLR)}.

\end{thebibliography}

\end{document}
